% !TEX root = ../main.tex
% Chapter 6

\chapter{Real-world results}

\label{Chapter6} % For referencing the chapter elsewhere, use~\ref{Chapter6}

\lhead{Chapter 6. \emph{Real-world results}} % This is for the header on each page - perhaps a shortened title

%----------------------------------------------------------------------------------------
\section{Outline}
\emph{Not intented for the reader.}
\begin{itemize}
  \item Apply proposed method to real-world applications, such as
    \begin{itemize}
      \item Daily life activity recognition (as the original context of this thesis is)
      \item PowerHouse sensor data
      \item Stock data?
    \end{itemize}
  \item Relate back to filter vs. wrapper methods - give results with different methods?
\end{itemize}

This chapter discusses the real-world data sets used for this research.
In \Cref{sec:data_sets} we discuss earlier used data sets and argue that these are inappropriate for our research.
It is followed by an overview of the setup we have created to record continuous data sets from an uncontrolled environment, both in- and outdoor.
The activities performed are discussed in \Cref{sec:data_gathering}.
As discussed in \Cref{Chapter4}, the algorithm is subject to user parameters.
In \Cref{sec:real_world_parameters} the effect and ranges of the parameters on the real world data sets are discussed.
Finally, the performance of the algorithm is discussed in \Cref{sec:real_world_performance}.

\input{./Chapters/Chapter6/data_sets}
\input{./Chapters/Chapter6/data_gathering}
\input{./Chapters/Chapter6/parameters}
\input{./Chapters/Chapter6/performance}
